---
title: "Effort_opport"
output: html_document
date: '2022-05-09'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Effort"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      cache = FALSE)

library(tidyverse)
library(lubridate)
library(rgdal)
library(sf)
library(oSCR)
library(mapview) 
library(raster)

theme_set(theme_light())
``` 

# Load data
In the data list you can find the detection data ("det") and traps ("traps") for each year between 2016 and 2019.   
For each detection there is :   
 - id: name of the bear which were detected  
 - month: the month of the detection (for us is the occasion). Here the month considered are between may and november
 - sex: the sex of the individual captured  
 - trap_id: the id I gave to the trap (f: match to traps in France, c: match to traps in Catalunya and op: match to the opportunistic grid at 5km)  
 - trap: id number of the trap to match with the trap file  
 - suivi: monitoring (systematic = structured sampling (hair snag, camera trap or both) and opportunsitic = indices samples after a depredation)  
      
For each trap:   
 - NOM: name of the itinerary where the trap is  
 - site: tells if the trap is composed of a hair snag alone (hair), a camera trap combine with hair snag (both), or an opportunistic trap define as centers of a 5km grid where at least one depredation event occurs since 2010.   
 - trap_id: the id I gave to the trap  
 - pays: country   
 - trap: id number of the trap  
 - suivi: monitoring (systematic = structured sampling (hair snag, camera trap or both) and opportunsitic = indices samples after a depredation)  
 - geometry: GPS location of the trap in UTM (crs= CRS("+proj=utm +zone=31 +datum=WGS84"))  
    
```{r}
load("data.RData")
```
   
In this case, stystematic/ structured monitoring and opportunistic
```{r}
# Encounters 
det_2017 <- data[["det"]][["2017"]] 

det_2018 <- data[["det"]][["2018"]]

det_2019 <- data[["det"]][["2019"]] 

# Traps positions
traps_2017 <- data[["traps"]][["2017"]] 

traps_2018 <- data[["traps"]][["2018"]] 

traps_2019 <- data[["traps"]][["2019"]] 
```
    
# I - Define edf and tdf
## 1) EDF : encounter data frame   
```{r}
edf <- data.frame(session = c(rep(1, nrow(det_2017)),
                              rep(2, nrow(det_2018)),
                              rep(3, nrow(det_2019))), # add the session column
                  ind = c(det_2017$id,
                          det_2018$id,
                          det_2019$id), # row names are individual IDs
                  occ = c(det_2017$month,
                          det_2018$month,
                          det_2019$month) %>% as.factor() %>% as.numeric, # occasions are month may = 1
                  trap = c(det_2017$trap,
                           det_2018$trap,
                           det_2019$trap), # col names are trap IDs
                  sex = fct_c(det_2017$sex, # note the use of fct_c !!
                              det_2018$sex,
                              det_2019$sex))

nocc <- 7 # number of occasion (month) per session (year) : may to november
```

## 2) TDF : trap data frame
The covariate "effort" indicate the number of visits per month in France. In may, june and september traps which are on itineraries are visited twice. 

###  A - TDF 2017 
```{r}
tdf2017 <- data.frame(trap = as.numeric(traps_2017$trap), # trap id number
                      X = st_coordinates(traps_2017)[,1], # longitude
                    Y = st_coordinates(traps_2017)[,2],# latitude
                      sep = rep("/",length(traps_2017$trap)), # separator
                      site = as.factor(traps_2017$site), # type of trap
                      pays = as.factor(traps_2017$pays), # country
                      suivi = as.factor(traps_2017$suivi)) # type of monitoring


# effort covariate : factor which account for the country and the number of visits per occasion in France
# 1 = one visit per month in France 
# 2 = two visits per month in France
# 3 = trap is in Spain
for (j in 1 : nocc) {
  eff <- rep(0, length(traps_2017$trap))
  for (i in 1:length(traps_2017$trap)){
    if (traps_2017$pays[i] == "Espagne"){ # if the trap is in Spain
      if(traps_2017$suivi[i] == "systematic"){
        eff[i] <- 3} # site systematic in Spain
      else {
        eff[i] <- 3} # site opportunist in Spain 
    }
    else {
      if (traps_2017$suivi[i] == "systematic"){
        if (traps_2017$effort[i] == "it" ){ # if the trap is on an transect/itinerary
          if (j == 1 | j == 2 | j == 5){ # and if we are in may, june or september
            eff[i] <- 2}  # the site is visited 2 times at this occasion
          else {
            eff[i] <- 1}} # the site is visited once
        else {
          eff[i] <- 1} # the site is visited once
      }
      else{
        eff[i] <- 1} # site opportunist in France 
    }
  } 
  name <- paste("effort", j, sep = ".")
  tdf2017[name] <- as.factor(eff)
}
```

###  B - TDF 2018  
```{r}
# Activity matrix
# Some itineraries begin in june : St Collombe, Escouloubre et Mijanes
notactiv <- traps_2018 %>% # identify traps which are on those itineraries 
  filter(NOM == "Mijanes1" | NOM == "Majanes2" | NOM == "Escouloubre" |NOM == "St Collombe")%>%
  st_drop_geometry()%>%
  dplyr::select(trap)%>%
  unlist()

trapactiv <- matrix(rep(1, length(traps_2018$trap)*7), ncol= 7) # initialisation of the activity matrix
trapactiv[notactiv,1] <- 0 # replace not active traps in june by 0

tdf2018 <- data.frame(trap = as.numeric(traps_2018$trap),
                      X = st_coordinates(traps_2018)[,1],
                      Y = st_coordinates(traps_2018)[,2],
                      TrapActiv = trapactiv,
                      sep = rep("/",length(traps_2018$trap)),
                      site = as.factor(traps_2018$site),
                      pays = as.factor(traps_2018$pays),
                      suivi = as.factor(traps_2018$suivi))

# effort covariate : factor which account for the country and the number of visits per occasion in France
for (j in 1 : nocc) {
  eff <- rep(0, length(traps_2018$trap))
  for (i in 1:length(traps_2018$trap)){
    if (traps_2018$pays[i] == "Espagne"){ # if the trap is in Spain
      if(traps_2018$suivi[i] == "systematic"){
        eff[i] <- 3} # site systematic in Spain
      else {
        eff[i] <- 3} # site opportunist in Spain
    }
    else {
      if (traps_2018$suivi[i] == "systematic"){
        if (traps_2018$effort[i] == "it" ){ # if the trap is on an transect
          if (j == 1 | j == 2 | j == 5){ # and if we are in may, june or september
            eff[i] <- 2}  # the site is visited 2 times at this occasion
          else {
            eff[i] <- 1}} # the site is visited once
        else {
          eff[i] <- 1} # the site is visited once
      }
      else{
        eff[i] <- 1} # site opportunist in France
    }
  } 
  name <- paste("effort", j, sep = ".")
  tdf2018[name] <- as.factor(eff)
}
```

###  C - TDF 2019  
```{r}
tdf2019 <- data.frame(trap = as.numeric(traps_2019$trap),
                      X = st_coordinates(traps_2019)[,1],
                      Y = st_coordinates(traps_2019)[,2],
                      sep = rep("/",length(traps_2019$trap)),
                      site = as.factor(traps_2019$site),
                      pays = as.factor(traps_2019$pays),
                      suivi = as.factor(traps_2019$suivi)) 

# effort covariate : factor which account for the country and the number of visits per occasion in France
for (j in 1 : nocc) {
  eff <- rep(0, length(traps_2019$trap))
  for (i in 1:length(traps_2019$trap)){
    if (traps_2019$pays[i] == "Espagne"){ # if the trap is in Spain
      if(traps_2019$suivi[i] == "systematic"){
        eff[i] <- 3} # site systematic in Spain
      else {
        eff[i] <- 3} # site opportunist in Spain
    }
    else {
      if (traps_2019$suivi[i] == "systematic"){
        if (traps_2019$effort[i] == "it" ){ # if the trap is on an transect
          if (j == 1 | j == 2 | j == 5){ # and if we are in may, june or september
            eff[i] <- 2}  # the site is visited 2 times at this occasion
          else {
            eff[i] <- 1}} # the site is visited once
        else {
          eff[i] <- 1} # the site is visited once
      }
      else{
        eff[i] <- 1} # site opportunist in France
    }
  } 
  name <- paste("effort", j, sep = ".")
  tdf2019[name] <- as.factor(eff)
}
```

# Save edf and tdf to run in oSCR (Ana)
```{r}
setwd("D:/MargSalas/Oso/Datos/Effort/France/Syst_Opport")
save(edf, file = "edf2017_2019_fr.RData")
save(tdf2017, file = "tdf2017_fr.RData")
save(tdf2018, file = "tdf2018_fr.RData")
save(tdf2019, file = "tdf2019_fr.RData")

```

# Vizualisation of traps 
```{r}
tdf2019 <- tdf2019 %>%
  st_as_sf(coords = c("X", "Y"), crs = CRS("+proj=utm +zone=31 +datum=WGS84"))
mapview(tdf2019, zcol = "effort.1")
```

